Viscous Transport in Eroding Porous Media

### 2019-07-30 docs/transport/erosion_transport.tex
started 20 body numerical results

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  Following the work of Ioakimidis et
al~\cite{ioa-pap-per1991} (see equation (2.12)), we mimick the
construction of the quadrature rule for $v(x)$ and $v'(x)$. We start
with the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\pi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x} dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivate
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
Similar to the quadrature rule for $v'(x)$ requring the value $v(x)$,
the quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$.
Both these quantities are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}.

@@

The first method to compute numerically compute the tortuosity uses $N_p$ trajectories
that are initially at $(1,y_i)$, $i=1,\ldots,N_p$, and stopped when
their $x$ coordinate is 1.  T

with originating at $x=-1$ and terminating at $x=1$, we use the approximation 
\begin{align}
  T = \frac{1}{d}\frac{\displaystyle\sum_{i=1}^{N_p} 
    u_x(y_i) \lambda(y_i) \Delta y_i}
  {\displaystyle\sum_{i=1}^{N_p} u_x(y_i) \Delta y_i}, 
\end{align}
where $y_i$ is the point on $S$ and $\Delta y_i = y_{i+1}-y_{i}$.
Similarly, to compute the tortuosity in terms of the Eulerian velocity
field, we use the approximation
\begin{align}
  T = \frac{\sum\limits_{i,j=1}^N \|\uu(\yy_{i,j})\|\Delta x_i \Delta y_j}
      {\sum\limits_{i,j=1}^N u(\yy_{i,j}) \Delta x_i\Delta y_j},
\end{align}
where $\Delta x_i = 1/N$

There have been efforts to relate the tortuosity to the
porosity~\cite{matyka2008tortuosity}, but these results can not be
universal since the porosity which depends on only a single number, the
tortuosity is affected by the distribution of pores within the porous
network.  In addition to the hydraulic tortuosity, a local tortuosity
can be used to analyze the local behavior of a streamline near a
body~\cite{matyka2008tortuosity}.  The local tortuosity of a streamline
initialized at $y$ along $S$ is
\begin{align*}
  \tau(y) = \frac{\lambda(y)}{d}.
\end{align*}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r1h_shear}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r0d5h_shear}
\includegraphics[width = 0.35 \textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{\label{fig:NearWall} 
  A single body eroding in Stokes flow. The color indicates the shear
  stress on the body. Therefore, erosion is fastest in the yellow
  regions and slowest in the blue regions.  In this simulation, we use
  three minimum body-to-wall initial distances $d=h$ (left), $h/2$
  (middle), and $h/10$ (right) with $N_\iin = 256$ discretization points
  on the body, a time step size of $\Dt=10^{-4}$, and smoothing
  parameters $\eps = 15/256$ and $\sigma=10/256$.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 nearly touching bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies a Hagen-Poiseuille flow.  The background
flow is 
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right].
\end{align}
Initially, $U=1$, and then $U$ is updated at each time step so that the
pressure drop from $x=-2$ to $x=2$ remains constant.  
{\color{red}
We
measure the area occupied by bodies 
during the eroding and record the flow rate $U$ quantifily in log scale
during the eroding in Fig.~\ref{fig:Eroding20flowrate}. 
Since we simulate the erosion by applying fixed pressure drop 
the flow rate $U$ is small when there are many bodies in the fluid and 
when the area fraction is getting smaller $U$ is getting larger.
In Fig.~\ref{fig:Eroding20flowrate}, 62.32\% of area of $[-1,1] \times
[-1,1]$ is occupied by 20 bodies initially and the flow rate $U$ starts from $9.86\times 10^{-5}$. 
Then the eroding speed is increasing
when more and more bodies are vanished in the fluid and 
the maximum eroding speed happens when the area fraction equals 7.57\% 
which is occupied by 8 bodies.
When all the bodies have vanished completely the flow rate $U$ is equal to 1.
In Fig.~\ref{fig:Eroding20flowrate}(b), the flow rate $U$ increases exponentially 
with growth rate close to $\ln10^4$. }

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/flow_rate20dense}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The area fraction of 20 nearly
 touching eroding bodies and the eroding speed versus time. 
(b) The incoming flow rate of $U$ in log scale versus time. 
Under fixed pressure drop condition, $U$ will be small initially,
increase when bodies are eroding, and grow exponentially 
to 1 when there is no body in the fluid. }
\end{figure}

@@

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense301}
%\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense1}
%\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense101}\\
%\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense201}
%\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} Erosion of 20 nearly touching bodies
  with the fixed pressure drop condition. The color is the vorticity of
  the fluid. The 4 snapshots are evenly spaced in time. We use $N_\iin =
  256$ discretization points on the body, a time-step of $\Dt=10^{-4}$,
  and smoothing parameters of $\eps= 15/256$ and $\sigma=10/256 $. In
  the forth frame, bodies 4, 12, and 17 have vanished completely, and
  body 19 has almost completely eroded.}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity study in the fluid with 
20 nearly touching bodies when the porosity is 62.9 \%.
(a) The 
x-component velocity u(-1, y) with respect to its maximum velocity $u_{max}=2.9828
\times 10^{-3}$ at equal spreading tracers on the cross section x = -1. 
(b) The local tortuosity $\tau(y)$ on the cross section x = -1. 
(c) The largest 10 difference of 
local tortuosity $\tau$ at two neighboring tracers in
 20 nearly touching bodies eroding when the porosity is 62.9 \% 
and the tortuosity is 1.0979. The same color of a pair of trajectories 
means they are from two neighboring tracers.}
%(c) The difference of $\tau(y)$ at two neighboring tracers on the cross section x = -1.
%(d) The ratio of the minimum to maximum speeds along the trajectories of tracers.}
\end{figure}

>>>>

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  Following the work of Ioakimidis et
al~\cite{ioa-pap-per1991} (see equation (2.12)), we mimick the
construction of the quadrature rule for $v(x)$ and $v'(x)$. We start
with the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\pi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x} dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivate
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
Similar to the quadrature rule for $v'(x)$ requring the value $v(x)$,
the quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$.
Both these quantities are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}.

@@

The first method to compute numerically compute the tortuosity uses
$N_p$ trajectories that are initially at $(1,y_i)$, $i=1,\ldots,N_p$,
and stopped when their $x$ coordinate is 1.  We choose $y_i$ so that the
trajectories are initially uniformly spaced in $(-1,1)$.  Having
computed these trajectories, the tortuosity is approximated as
\begin{align}
  T = \frac{1}{d}\frac{\displaystyle\sum_{i=1}^{N_p} 
    u_x(y_i) \lambda(y_i) \Delta y}
  {\displaystyle\sum_{i=1}^{N_p} u_x(y_i) \Delta y_i}, 
\end{align}
where $\Delta y = 2/(N_p + 1)$ and $y_i = -1 + i \Delta y$,
$i=1,\ldots,N_p$.  Alternatively, to compute the tortuosity in terms of
the Eulerian velocity field, we first compute the velocity at $(x_i,y_j)
= (-1 + i\Delta x, -1 + j\Delta y)$, $i,j=1,\ldots,N$, where $\Delta x =
\Delta y = 2/N$, and the velocity at points inside an eroding body are
assigned a value of 0.  Then, the tortuosity is approximated as
\begin{align}
  T = \frac{\sum\limits_{i,j=1}^N \|\uu(x_i,y_j)\|\Delta x \Delta y}
      {\sum\limits_{i,j=1}^N u(x_i,y_j) \Delta x \Delta y},
\end{align}
where $\uu = (u,v)$.

There have been efforts to relate the tortuosity to the
porosity~\cite{matyka2008tortuosity}, but these results can not be
universal since the porosity which depends on only a single number, the
tortuosity is affected by the distribution of pores within the porous
network.  In addition to the hydraulic tortuosity, a local tortuosity
can be used to analyze the local behavior of a streamline near a
body~\cite{matyka2008tortuosity}.  The local tortuosity of a streamline
initialized at $y$ along $S$ is
\begin{align*}
  \tau(y) = \frac{\lambda(y)}{d}.
\end{align*}
\todo[inline]{Need to flesh this out more}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r1h_shear}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r0d5h_shear}
\includegraphics[width = 0.38 \textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{\label{fig:NearWall} 
  A single body eroding in Stokes flow. The color indicates the shear
  stress on the body. Therefore, erosion is fastest in the yellow
  regions and slowest in the blue regions.  In this simulation, we use
  three minimum body-to-wall initial distances $d=h$ (left), $h/2$
  (middle), and $h/10$ (right) with $N_\iin = 256$ discretization points
  on the body, a time step size of $\Dt=10^{-4}$, and smoothing
  parameters $\eps = 15/256$ and $\sigma=10/256$.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 nearly touching bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies initially placed in $[-1,1] \times
[-1,1]$.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 2.  For equispaced instances of the
configuration are plotted in Figure~\ref{fig:Eroding20vort}.  The color
in these plots is the vorticity, and this can be used to determine
regions of high and low erosion---erosion happens fastest along regions
of the eroding bodies with a large vorticity in magnitude.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  Initially, 62\%  of the geometry eroded
bodies corresponding to a porosity of $0.38$.  As the bodies erode, the
area fraction decreases towards 0.  The trend of the area fraction
resembles that of our previous work~\cite{qua-moo2018} (Figure 10(a)),
but with a larger initial area fraction.  We also compute the flow rate
$U$ required to maintain a constant pressure drop across the channel in
Figure~\ref{fig:Eroding20flowrate}(b).  Again, the trend of $U$
resembles that of our previous work~\cite{qua-moo2018} (Figure 10(b)),
except that the initial flow rate is an order of magnitude smaller
because of the larger initial area fraction.

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} Erosion of 20 nearly touching bodies
with the fixed pressure drop condition. The color is the vorticity of
the fluid. The 4 snapshots are evenly spaced in time. We use $N_\iin =
256$ discretization points on the body, a time-step of $\Dt=10^{-4}$,
and smoothing parameters of $\eps= 15/256$ and $\sigma=10/256 $. In the
forth frame, bodies 4, 12, and 17 have vanished completely, and body 19
has almost completely eroded.}
\end{center}
\end{figure}

{\color{red}
We
measure the area occupied by bodies 
during the eroding and record the flow rate $U$ quantifily in log scale
during the eroding in Fig.~\ref{fig:Eroding20flowrate}. 
Since we simulate the erosion by applying fixed pressure drop 
the flow rate $U$ is small when there are many bodies in the fluid and 
when the area fraction is getting smaller $U$ is getting larger.
In Fig.~\ref{fig:Eroding20flowrate}, 62.32\% of area of $[-1,1] \times
[-1,1]$ is occupied by 20 bodies initially and the flow rate $U$ starts from $9.86\times 10^{-5}$. 
Then the eroding speed is increasing
when more and more bodies are vanished in the fluid and 
the maximum eroding speed happens when the area fraction equals 7.57\% 
which is occupied by 8 bodies.
When all the bodies have vanished completely the flow rate $U$ is equal to 1.
In Fig.~\ref{fig:Eroding20flowrate}(b), the flow rate $U$ increases exponentially 
with growth rate close to $\ln10^4$. }

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/porosity20dense}
\todo[inline]{SH: Make the y-limits of this figure go from 0 to 0.7 ...
no need to make it go below 0}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/flow_rate20dense}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The area fraction of 20 nearly
 touching eroding bodies and the eroding speed versus time. 
(b) The incoming flow rate of $U$ in log scale versus time. 
Under fixed pressure drop condition, $U$ will be small initially,
increase when bodies are eroding, and grow exponentially 
to 1 when there is no body in the fluid. }
\end{figure}

@@

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity study in the fluid with 
20 nearly touching bodies when the porosity is 62.9 \%.
(a) The 
x-component velocity u(-1, y) with respect to its maximum velocity
$u_{\max}=2.9828
\times 10^{-3}$ at equal spreading tracers on the cross section x = -1. 
(b) The local tortuosity $\tau(y)$ on the cross section x = -1. 
(c) The largest 10 difference of 
local tortuosity $\tau$ at two neighboring tracers in
 20 nearly touching bodies eroding when the porosity is 62.9 \% 
and the tortuosity is 1.0979. The same color of a pair of trajectories 
means they are from two neighboring tracers.}
%(c) The difference of $\tau(y)$ at two neighboring tracers on the cross section x = -1.
%(d) The ratio of the minimum to maximum speeds along the trajectories of tracers.}
\end{figure}
